{\large Master Thesis}\\{\normalsize on

### 2011-07-20 doc/report.tex
m

Different algorithms exists for solving the RMQ problem, and we can solve it in optimal $O(1)$ time and $O(n)$ space. But like for the LCE problem, the asymptotically best RMQ solution may not be the best RMQ solution in practice. We examine the following RMQ solutions, where \RMQpq{p}{q} has $O(p)$ space and preprocessing time, and $O(q)$ query time:
\begin{samepage}
\begin{description}
\item[\RMQn] walks through the array directly in $O(n)$ time and does not use any preprocessing (except for transforming LCE into RMQ).
\item[\RMQq] preprocesses the array in $O(n)$ space in order to do queries in $O(1)$ time. This is a twolevel data structure as shown in \pref{fig:sa+lcp+min-twolevel}. The array is split into blocks of size $O(\log n)$. The top level is an array of length $O(n/\log n)$, where each element contains the minimum value of one block. The top level uses a \RMQpq{n\log n}{1}-solution, and the lower level precomputes all answers for each block, which is done in $O(n)$ space, since there are only a limited number of different block types of size $O(\log n)$.
\item[\RMQlog] is the same as \RMQq\ except that the lower level is replaced with a direct search through a $O(\log n)$ length block.
\end{description}
\end{samepage}

@@

\begin{proof}
We first construct the SA and LCP arrays, which takes $O(\sortt)$ time~\cite{sort-complexity}. We then preprocess each of the $k$ levels in $O(n)$ time, since we loop through $n$ substrings, and comparing neighboring substrings takes constant time when we use the LCE array. The total preprocessing time becomes $O(\sortt + k\cdot n)$.
\end{proof}

@@

The amount if I/O's used by \fprintk\ is $O(k\cdot n^{1/k})$. However if we structure our tables of fingerprints differently, we can improve the number of I/O's to $O(k(n^{1/k}/B+1))$ in the cache-oblivious model. Instead of storing the fingerprint of $s[i\twodots i+t_\ell-1]$ at $H_\ell[i]$, we can store it at $H_\ell[((i-1)\mod t_\ell)\cdot\lceil n/t_\ell\rceil+\lfloor (i-1)/t_\ell\rfloor+1]$. This will group all used fingerprints at level $\ell$ next to each other in memory, such that the amount of I/O's at each level is reduced from $O(n^{1/k})$ to $O(n^{1/k}/B)$.

@@

The algorithm keeps the invariant $\zlabel(p_i[0\twodots v]) = \zlabel(p_j[0\twodots v])$. At each iteration we find $v'$, which is the length of the longest common prefix of $\zlabel(p_i[v\twodots v+m_i])$ and $\zlabel(p_j[v\twodots v+m_j])$. This gives $\zlabel(p_i[0\twodots v+v']) = \zlabel(p_j[0\twodots v+v'])$, and we therefore keep the invariant when we add $v'$ to $v$ in step five.

@@

\begin{lemma}
We can preprocess the \proc{LadderLce} data structure in $O(\sortt)$ time.
\end{lemma}
\begin{proof}
Constructing $s$ from the ladders takes $O(n)$ time, and preprocessing it for constant time LCE takes $O(\sortt)$ time. The rest of the data structures have $O(n)$ preprocessing times.
\end{proof}

>>>>

Different algorithms exists for solving the RMQ problem, and we can solve it in optimal $O(1)$ time and $O(n)$ space. But like for the LCE problem, the asymp-\\totically best RMQ solution may not be the best RMQ solution in practice. We examine the following RMQ solutions, where \RMQpq{p}{q} has $O(p)$ space and preprocessing time, and $O(q)$ query time:
\begin{samepage}
\begin{description}
\item[\RMQn] walks through the array directly in $O(n)$ time and does not use any preprocessing (except for transforming LCE into RMQ).
\item[\RMQq] preprocesses the array in $O(n)$ space in order to do queries in $O(1)$ time. This is a twolevel data structure as shown in \pref{fig:sa+lcp+min-twolevel}. The array is split into blocks of size $O(\log n)$. The top level is an array of length $O(n/\log n)$, where each element contains the minimum value of one block. The top level uses a \RMQpq{n\log n}{1}-solution, and the lower level precomputes all answers for each block, which is done in $O(n)$ space, since there are only a limited number of different block types of size $O(\log n)$.
\item[\RMQlog] is the same as \RMQq\ except that the lower level is replaced with a direct search through a $O(\log n)$ length block.
\end{description}
\end{samepage}

@@

\begin{proof}
We first construct the SA and LCP arrays, which takes $O(\sortt)$ time~\cite{sort-complexity}. We then preprocess each of the $k$ levels in $O(n)$ time, since we loop through $n$ substrings, and comparing neighboring substrings takes constant time when we use the LCE array. The total preprocessing time becomes $O(\sortt\\ + k\cdot n)$.
\end{proof}

@@

The amount of I/O's used by \fprintk\ is $O(k\cdot n^{1/k})$. However if we structure our tables of fingerprints differently, we can improve the number of I/O's to $O(k(n^{1/k}/B+1))$ in the cache-oblivious model. Instead of storing the fingerprint of $s[i\twodots i+t_\ell-1]$ at $H_\ell[i]$, we can store it at $H_\ell[((i-1)\mod t_\ell)\cdot\lceil n/t_\ell\rceil+\lfloor (i-1)/t_\ell\rfloor+1]$. This will group all used fingerprints at level $\ell$ next to each other in memory, such that the amount of I/O's at each level is reduced from $O(n^{1/k})$ to $O(n^{1/k}/B)$.

@@

The algorithm keeps the invariant $\zlabel(p_i[0\twodots v]) = \zlabel(p_j[0\twodots v])$. At each iteration we find $v'$, which is the length of the longest common prefix of $\zlabel(p_i[v\twodots v+m_i])$ and $\zlabel(p_j[v\twodots v+m_j])$. This gives $\zlabel(p_i[0\twodots v+v']) = \\\zlabel(p_j[0\twodots v+v'])$, and we therefore keep the invariant when we add $v'$ to $v$ in step five.

@@

\begin{lemma}
We can preprocess the \proc{LadderLce} data structure in $O(\sortt)$ time.
\end{lemma}
\begin{proof}
Constructing $s$ from the ladders takes $O(n)$ time, and preprocessing it for constant time LCE takes $O(\sortt)$ time. The rest of the data structures have\\ $O(n)$ preprocessing times.
\end{proof}
